\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{amsfonts,amsmath}
\usepackage{graphicx,float}
% -----------------------------------
\usepackage{hyperref}
\hypersetup{%
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue,
  linkbordercolor={0 0 1}
}
% -----------------------------------
\usepackage{fancyhdr}
\newcommand\course{MATH-UA.0252\\Numerical Analysis}
\newcommand\hwnumber{12}                  % <-- homework number
\newcommand\NetIDa{Ryan Sh\`iji\'e D\`u} 
\newcommand\NetIDb{December 9th, 2022}
\pagestyle{fancyplain}
\headheight 35pt
\lhead{\NetIDa\\\NetIDb}
\chead{\textbf{\Large Worksheet \hwnumber}}
\rhead{\course}
\lfoot{}
\cfoot{}
\rfoot{\small\thepage}
\headsep 1.5em
% -----------------------------------
\usepackage{titlesec}
\renewcommand\thesubsection{(\arabic{section}.\alph{subsection})}
\titleformat{\subsection}[runin]
        {\normalfont\bfseries}
        {\thesubsection}% the label and number
        {0.5em}% space between label/number and subsection title
        {}% formatting commands applied just to subsection title
        []% punctuation or other commands following subsection title
% -----------------------------------
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}
% -----------------------------------
\input{../../command.tex}
\begin{document}

\section{Convergence order of quadrature}
\subsection{}
We would like to integrate a function on $[0,1]$ using the composite trapezoid rule with sub-interval size $h$. We name the result $T_h$. How does the error ($e_h = |T_n-I|$) scale with $h$?

\subsection{}
In real application, we do not know the true value of the integral. To verify the order of convergence of our code, we can calculate this quantity:
\begin{align*}
    \frac{T_h-T_{h/2}}{T_{h/2}-T_{h/4}}.
\end{align*}
How does this quantity scale with $h$?

\subsection{}
We will show this using the code from last session.

\section{Inner product}
\subsection{}
Let's define a new inner product in $\mathbb R^3$ with a symmetric
  positive-definite matrix $W\in \mathbb R^{3\times 3}$ as follows:
  $$
  \langle \boldsymbol u, \boldsymbol v\rangle_W := \boldsymbol u^\top W \boldsymbol v
  $$
  Show that this defines an inner product on $\mathbb R^3$ and write
  down the induced norm. If 
  $$W=\begin{bmatrix}2& -1& 0 \\ -1 & 2 & -1
  \\ 0 & -1 & 2 \end{bmatrix},$$
  give a vector that is orthogonal to
  $[1,0,0]^\top$ in the $W$-inner product.

\subsection{}
Now let's consider a symmetric positive
  semi-definite matrix. Does the above definition still define an
  inner product and why/why not?
  
\section{Orthogonal Polynomial}
We assume
  an inner product on $[-1,1]$ with weight $\omega(x):=1-x^2$, i.e.,
  $$
  \langle p,q\rangle = \int_{-1}^1 \omega(x)p(x)q(x)\,\de x.
  $$ We define the following polynomials, which are orthogonal with
  respect to this inner product:
  $$
  \varphi_0(x) = 1, \quad \varphi_1(x) = 2x, \quad \varphi_2(x) = 5x^2-1
  $$

\subsection{}
Verify that $\varphi_0$ and $\varphi_1$ are orthogonal on $[-1,1]$
    with respect to the weight $\omega$.

\subsection{}
Are the $\varphi_j$'s orthonormal under this inner product?

\subsection{}
Find the polynomial $p \in \boldsymbol P_2$ for which
    $\displaystyle \int_{-1}^1 (1-x^2)(p(x) - x^3)^2 \, \de x$ is minimal.




% \vfill
% \bibliographystyle{alpha}
% \bibliography{citation}

\end{document}