\documentclass[11pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{amsfonts,amsmath}
\usepackage{graphicx,float}
% -----------------------------------
\usepackage{hyperref}
\hypersetup{%
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue,
  linkbordercolor={0 0 1}
}
% -----------------------------------
\usepackage{fancyhdr}
\newcommand\course{MATH-UA.0252\\Numerical Analysis}
\newcommand\hwnumber{3-1}                  % <-- homework number
\newcommand\NetIDa{Ryan Sh\`iji\'e D\`u} 
\newcommand\NetIDb{September 29th, 2023}
\pagestyle{fancyplain}
\headheight 35pt
\lhead{\NetIDa\\\NetIDb}
\chead{\textbf{\Large Worksheet \hwnumber}}
\rhead{\course}
\lfoot{}
\cfoot{}
\rfoot{\small\thepage}
\headsep 1.5em
% -----------------------------------
\usepackage{titlesec}
\renewcommand\thesubsection{(\arabic{section}.\alph{subsection})}
\titleformat{\subsection}[runin]
        {\normalfont\bfseries}
        {\thesubsection}% the label and number
        {0.5em}% space between label/number and subsection title
        {}% formatting commands applied just to subsection title
        []% punctuation or other commands following subsection title
% -----------------------------------
\setlength{\parindent}{0.0in}
\setlength{\parskip}{0.1in}
% -----------------------------------
\input{../../command.tex}
\begin{document} 
\section{Complexity of computing matrix condition numbers}
\subsection{}\label{eq:comp_norm_a}
Recall the formulas for the 1- and $\infty$-norm. Assuming that taking the absolute value and determining the maximum does not contribute to the overall computational cost, calculate how many \textit{flops} (floating point operations) are needed to calculate $\norm{A}_1$ and $\norm{A}_\infty$ for $A\in\mathbb{R}^{n\times n}$? By what factor will the calculation time increase when you double the matrix size?

\subsection{}\label{eq:comp_norm_b}
Now implement a simple code that calculates $\norm{A}_1$ and $\norm{A}_\infty$ for a matrix of any size $n\geq 1$. Try to do this without using loops\footnote{The commands needed in Python \texttt{np.abs} and \texttt{np.sum}. Most commands can not only applied to numbers, but also to vectors, where they apply to each component; this is typically much faster.}! Using system sizes of $n_1=100$, $n_{k+1}=2n_k,k=1,\ldots,7$, determine how long your code takes to calculate $\norm{A}_1$ and $\norm{A}_\infty$ for a matrix $A\in\mathbb{R}^{n_i\times n_i}$ with random entries and report the results. Can you confirm the estimate from \ref{eq:comp_norm_a}?

\subsection{}
Python has the built-in functions \texttt{np.linalg.norm} to calculate matrix norms. Calculate for the system sizes in \ref{eq:comp_norm_b} $\norm{A}_1$ and $\norm{A}_\infty$ using both your implementation and the built-in \texttt{norm} function, determine for each $n_i$ how long each code takes and plot the results in one graph. On average, by what factor is MATLAB or Python's implementation faster than yours?



\section{Condition numbers based on different norms}
\subsection{}
Let $A \in \mathbb{R}^{n\times n}$ be defined by
\begin{equation*}
    A = %
    \begin{bmatrix}
      1      & 0      & 0      & 0      &  0      \\ 
      1      & 1      & 0      & 0      &  0      \\
      1      & 0      & 1      & 0      &  0      \\
      \vdots & \vdots & \vdots & \ddots &  \vdots \\
      1      & 0      & 0      & 0      &  1
    \end{bmatrix}.
\end{equation*}
Calculate $\kappa_1(A)$ and $\kappa_\infty(A)$. We see that a matrix can be well or ill-conditioned depending on the choice of norms.

\subsection{}
Indeed, we solve $A\ve x = \ve b$ and $A(\ve x+\Delta \ve x) = (\ve b+\Delta \ve b)$ where
\begin{align*}
    \ve b = \begin{bmatrix}
    1\\1\\\vdots\\1
    \end{bmatrix},\qquad \ve x = \begin{bmatrix}
    1\\0\\\vdots\\0
    \end{bmatrix},\qdt{and} \Delta\ve b = \begin{bmatrix}
    \epsilon\\0\\\vdots\\0
    \end{bmatrix}
\end{align*}
Check that we have for both norms:
\begin{align*}
    \frac{\norm{\Delta\ve x}}{\norm{\ve x}} \leq \kappa(A)\frac{\norm{\Delta\ve b}}{\norm{\ve b}}.
\end{align*}

\section{Conditional number for the Hilbert matrix}  
The Hilbert matrix $H\in \mathbb R^{n\times n}$ is a matrix with
  entries
  $$
  h_{ij} = \frac{1}{i+j-1}.
  $$ 
\subsection{}
Using MATLAB or Python, compute the 2-norm-based condition
  numbers for $n=3,5,10,20,25$.  
  
\subsection{}
  Let's consider a relative right hand
  side perturbation $\delta\boldsymbol b$ of a linear system with
  $\|\delta\boldsymbol b\|_2/\|\boldsymbol b\|_2\approx
  10^{-15}$. Write down the corresponding bounds $\|\delta\boldsymbol
  x\|_2/\|\boldsymbol x\|_2$ from the theory we discussed in class.

\subsection{}
  Now, let's compute the actual error. Use the right-hand side vector
  with entries $b_i = \sum_{j=1}^n(j/(i+j-1))$ chosen such that the
  solution vector has entries $x_i=i$. Now, Compute the numerical
  solutions\footnote{Note that all these computations contain tiny
    errors due to the final precision of computer computations.}
  $\boldsymbol x$, then re-compute $\boldsymbol b=H\boldsymbol x$ and
  compare the relative right-hand side error and the relative error
  in the solutions. How much are these better than the estimates you
  got from the condition number?
    
    
% \bibliographystyle{alpha}
% \bibliography{citation}


\end{document}